---
title: "Midterm  Project: Hedonic Home Price Prediction"
author: "Emma Sun"
date: "10/15/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

## I. Introduction & Motivation


```{r Load Libraries, echo=TRUE, message=FALSE, warning=FALSE}
# 1. load Libraries
library(sf)
library(tidyverse)
# install.packages('mapview')
library(mapview)
library(spdep)
library(caret)
library(ckanr) # for opening data APIs built on CKAN technology
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)
library(jtools)     # for regression model plots
library(stargazer) # for creating table
library(broom)
library(tufte)
library(rmarkdown)
library(kableExtra)
library(tidycensus)

# 2. Identify functions
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 15,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.text.x = element_text(size = 14))
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(color = "darkred", size=15, face="bold"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}

#3. Load Quantile break functions

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}

q5 <- function(variable) {as.factor(ntile(variable, 5))}


# Load hexadecimal color palette

palette <- c('#feedde', '#fdbe85', '#fd8d3c', '#e6550d', '#a63603')

# for calculating average nearest neighbor distance.

nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    dplyr::summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull() # pull() is similar to $. It's mostly useful because it looks a little nicer in pipes, it also works with remote data frames, and it can optionally name the output.
  
  return(output)  
}

```

## II. Data Manipulation and Visualization
### 2.0 Setup
In this section, we loaded necessary libraries, created plot theme options and map theme options, and identified functions of quintile breaks, and average nearest neighbor distance for further analysis.

### 2.1 Data Wrangling 

In this section, we loaded Miami data and other complementary datasets, conducted feature selection and engineering. Besides, for other analysis, we separate the whole dataset into `miami.training` and `miami.test` in advance.  


#### 2.1.1 Data loading

Five categorical  datasets were used in this project:

**Miami House Price and Internal Characteristics**: basic geometric dataset provided by the course in advance.

**Miami Base Data**: geometric data containing the geographic information of Miami and Miami Beach. 

**Miami Neighborhood Shapefiles**: digital geographic [dataset](https://hub.arcgis.com/datasets/2f54a0cbd67046f2bd100fb735176e6c_0) contains polygon features representing neighborhood areas within the City of Miami as well as Miami Beach, Florida. 

**Census Data**: demographic variables from the ACS 2017 for census tracts in Miami-Dade County

**Amenity Data**: compilatory data chosen for describing amenities and public service of Miami-Dade County. The datasets we chose encompasses crime rate, education opportunity, healthcare service and entertainment accessibility. See each dataset below:

1. [**Sexual Predator**](https://gis-mdc.opendata.arcgis.com/datasets/sexual-predator/data): A point feature class of Sexual Offenders and Predators within Miami-Dade County that is used by the Sexual Offender/Predator Residence Search Internet application.

2. [**Public School**](https://gis-mdc.opendata.arcgis.com/datasets/public-school): A point feature class of the Miami-Dade County Public Schools facilities.

3. [**College, University**](https://gis-mdc.opendata.arcgis.com/datasets/college): A point feature class of colleges and universities within Miami-Dade County.

4. [**Hospital**](https://gis-mdc.opendata.arcgis.com/datasets/hospital-1): A point feature class of the Hospital facilities within Miami-Dade County.

5. [**Landmark**](https://gis-mdc.opendata.arcgis.com/datasets/landmark): A point feature class of landmarks within Miami-Dade County.

6. [**Hotel, Motel, Inn**](https://gis-mdc.opendata.arcgis.com/datasets/hotel-motel-inn): A point feature class of Hotel, Motel and Inn facilities within Miami-Dade County.

7. [**Shopping Mall**](https://gis-mdc.opendata.arcgis.com/datasets/major-mall): A point feature class that contains the locations of major shopping malls within Miami-Dade County. Shopping Mall is considered a large retail complex containing a variety of stores and often restaurants in which one or more buildings form a complex of shops representing merchandisers with interconnecting walkways that enable customers to walk from unit to unit.


```{r Base, message=FALSE, warning=FALSE, results = FALSE}

# I. Base data
## House Price & internal characteristics
miami.sf <- st_read('C:/Users/Yiming Ma/Documents/UPenn/MUSA508/hw/hw2/studentsData.geojson')

#miami.sf <- st_read('/Users/penguin/Box #Sync/GitHub/MUSA508-Midterm/studentsData.geojson')

#miami.sf <- st_read('/Users/penguin/Box #Sync/GitHub/MUSA508-Midterm/studentsData.geojson')

## Miami base data
miami.base <-st_read("https://opendata.arcgis.com/datasets/5ece0745e24b4617a49f2e098df8117f_0.geojson") %>%
  filter(NAME %in% c("MIAMI", "MIAMI BEACH"))

## Neighborhood data
miami.neigh <- st_read('https://opendata.arcgis.com/datasets/2f54a0cbd67046f2bd100fb735176e6c_0.geojson')
# plot(miami.neigh)

miami.sf <-st_join(miami.sf,miami.neigh)

### check neighborhood information
miami.neigh.num <- as.data.frame(table(miami.sf$LABEL)) 

## Census Tract data
## View(load_variables(2017,'acs5',cache = TRUE))
tracts17 <- 
  get_acs(geography = "tract", variables = c("B25026_001E","B02001_002E",
                                             "B19013_001E","B25058_001E",
                                             "B06012_002E", 
                                             # vacant variables
                                             "B25002_003E", 
                                             "B25004_002E","B25004_003E",
                                             "B25004_004E","B25004_005E",
                                             # total housing unit
                                             "B25001_001E",
                                             # renter occupied 
                                             'B08137_003E'), 
          year=2017, state= 12, county= 086, geometry=T, output="wide") %>%
  st_transform('EPSG:2236') %>%
  rename(TotalPop = B25026_001E, 
         Whites = B02001_002E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E,
         TotalVacant = B25002_003E,
         ForRent = B25004_002E,
         ForRentVac = B25004_003E,
         ForSale = B25004_004E,
         ForSaleVac = B25004_005E,
         TotalUnit = B25001_001E,
         RenterOccupied = B08137_003E
         ) %>%
  dplyr::select(-NAME, -starts_with("B")) %>% #-starts_with("B") awesome!
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         pctTotalVacant = ifelse(TotalUnit > 0, TotalVacant / TotalUnit, 0),
         TotalOccupied = TotalUnit - TotalVacant,
         pctRenterOccupied = ifelse(TotalOccupied >0, RenterOccupied/TotalOccupied, 0)) %>%
  dplyr::select(-Whites, -TotalPoverty) 


projected.tracts17 <- 
  tracts17 %>% 
  st_transform(st_crs(miami.sf))

miami.sf <-st_join(miami.sf, projected.tracts17)


## Amenity Data
## 1. Miami landmark
miami.landmark <- st_read('https://opendata.arcgis.com/datasets/70a14825e66f4f0eb28d2a9cceba1761_0.geojson')
miami.landmark.sf <- miami.landmark %>%
  select(geometry) %>%
  na.omit() %>%
  distinct()

## 2. Miami Shopping Mall
miami.mall <- st_read('https://opendata.arcgis.com/datasets/cb24d578246647a9a4c57bbd80c1caa8_0.geojson')
miami.mall.sf <- miami.mall %>%
  select(geometry) %>%
  na.omit() %>%
  distinct()

## 3. Miami Sexual Predator
miami.sexual <- st_read('https://opendata.arcgis.com/datasets/f8759d722aeb4198bfe7c4ad780604d2_0.geojson')

miami.sexual.sf <- miami.sexual %>%
  select(geometry) %>%
  na.omit() %>%
  distinct()

## 4. School data
miami.school <- st_read('https://opendata.arcgis.com/datasets/d3db0fce650d4e40a5949b0acae6fe3a_0.geojson')
miami.school.sf <- miami.school %>%
  select(geometry) %>%
  na.omit() %>%
  distinct()


## 5. Hotel data
miami.hotel <- st_read('https://opendata.arcgis.com/datasets/d37bbc15e7304b4ca4607783283147b7_0.geojson')
miami.hotel.sf <- miami.hotel%>%
  select(geometry) %>%
  na.omit() %>%
  distinct()

## 6. College data
miami.college <- st_read('https://opendata.arcgis.com/datasets/7db056c406b943dc8f3f377b99d77588_0.geojson')
miami.college.sf <- miami.college%>%
  select(geometry) %>%
  na.omit() %>%
  distinct()

## 7. Hospital data
miami.hospital <- st_read('https://opendata.arcgis.com/datasets/0067a0e8b40644f980afa23ad34c32c4_0.geojson')
miami.hospital.sf <- miami.hospital %>%
  select(geometry) %>%
  na.omit() %>%
  distinct()

```

#### 2.1.2 Feature Engineering

In this section, we created some new features to describe the amenities and public services in Miami-Dade County, as well as to each home sale observation’s internal characteristics.

First of all, we were aimed at creating features to measure the extent to which buyers can assess multiple public services, including hospitals, shopping malls, landmarks, hotels, motels and inns, and education resources. To utmost avoid scale bias triggered by setting arbitrary areal unit or fixed buffer distance of each home sale observation,  we chose to calculate the average nearest neighbor distance from each home sale to its *k* nearest neighbor public services. Accordingly, the smaller the value is, the more possibility that the observation can access specific service. To better compare each feature’s impact, we set the *k* from 3 to 5.

By leveraging the exactly same approach, we also created a new feature to measure the extent to which buyers can expose sexual assaults. 

Secondly, we added four more features to indicate the duration each house was built, effectively used, and if the house was built with a pool or a patio.


```{r Feature Engineering, message=FALSE, warning=FALSE, results = FALSE}
## 1. Landmark
miami.sf <-
  miami.sf %>% 
  mutate(
    landmark_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.landmark.sf)), 1),
    landmark_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.landmark.sf)), 2), 
    landmark_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.landmark.sf)), 3), 
    landmark_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.landmark.sf)), 4), 
    landmark_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.landmark.sf)), 5)) 


## 2. Shopping mall
miami.sf <-
  miami.sf %>% 
  mutate(
    mall_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.mall.sf)), 1),
    mall_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.mall.sf)), 2), 
    mall_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.mall.sf)), 3), 
    mall_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.mall.sf)), 4), 
    mall_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.mall.sf)), 5)) 

## 3. Sexual Assaults
miami.sf <-
  miami.sf %>% 
  mutate(
    sexual_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.sexual.sf)), 1),
    sexual_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.sexual.sf)), 2), 
    sexual_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.sexual.sf)), 3), 
    sexual_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.sexual.sf)), 4), 
    sexual_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.sexual.sf)), 5)) 


## 4. Hotel Access
miami.sf <-
  miami.sf %>% 
  mutate(
    hotel_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hotel.sf)), 1),
    hotel_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hotel.sf)), 2), 
    hotel_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hotel.sf)), 3), 
    hotel_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hotel.sf)), 4), 
    hotel_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hotel.sf)), 5)) 


## 5. School Access

miami.sf <-
  miami.sf %>% 
  mutate(
    school_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.school.sf)), 1),
    school_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.school.sf)), 2), 
    school_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.school.sf)), 3), 
    school_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.school.sf)), 4), 
    school_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.school.sf)), 5)) 


## 6. College Access
miami.sf <-
  miami.sf %>% 
  mutate(
    college_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.college.sf)), 1),
    college_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.college.sf)), 2), 
    college_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.college.sf)), 3), 
    college_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.college.sf)), 4), 
    college_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.college.sf)), 5)) 

## 7. Hospital Access
miami.sf <-
  miami.sf %>% 
  mutate(
    hospital_nn1 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hospital.sf)), 1),
    hospital_nn2 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hospital.sf)), 2), 
    hospital_nn3 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hospital.sf)), 3), 
    hospital_nn4 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hospital.sf)), 4), 
    hospital_nn5 = nn_function(st_coordinates(st_centroid(miami.sf)), st_coordinates(st_centroid(miami.hospital.sf)), 5)) 

## 8. Internal characteristics
miami.sf <-
  miami.sf %>% 
  mutate(pool_house = ifelse(str_detect(XF1, "Pool"), 1, 0),
         patio_house = ifelse(str_detect(XF1, "Patio"), 1, 0),
         Age = 2020- YearBuilt,
         Age.effective = 2020- EffectiveYearBuilt)

```


#### 2.1.3 Split dataset by 'toPredict'

As indicated earlier, we separated the whole dataset into `miami.training` and `miami.test` in advance for model building in next section. 

```{r Split Dataset, echo=TRUE, message=FALSE, warning=FALSE}
# Split dataset 
miami.training <- miami.sf %>%
  filter(toPredict == '0')

# summary(miami.training)

miami.test <- miami.sf %>%
  filter(toPredict == '1') 

# summary(miami.test)
```


### 2.2 Summary Statistics of Features

```{r All Features, echo=TRUE, message=FALSE, warning=FALSE}
## All potential features
all.feature.list <- miami.training %>%
  dplyr:: select(-saleDate,-saleType,-saleQual,-saleYear,-Property.Address,
         -Year,-WVDB,-HEX,-GPAR,-County.2nd.HEX,-County.Senior,-County.LongTermSenior,
         -County.Other.Exempt,-City.2nd.HEX,-City.Senior,-City.LongTermSenior,
         -City.Other.Exempt,-MillCode,-Land.Use,-Owner1,-Owner2,-Mailing.Address,
         -Mailing.City,-Mailing.State,-Mailing.Zip,-Mailing.Country,
         -starts_with("Legal"), -YearBuilt, -EffectiveYearBuilt, -X, -FID, -starts_with("Shape_"), -Folio,
         -XF1, -XF2, -XF3) %>%
  st_drop_geometry()




```

```{r Three, echo=TRUE, message=FALSE, warning=FALSE, results ='asis'}
## 2.1 Feature: Internal Characteristics
internal.feature.list <- all.feature.list %>%
  dplyr:: select(SalePrice, Land, Bldg, Total, Assessed,County.Taxable,City.Taxable,AdjustedSqFt,
                 LotSize, Bed, Bath, Stories, Units, LivingSqFt, ActualSqFt, Age,Age.effective,
                 pool_house, patio_house) 

stargazer(internal.feature.list, type = "html", 
          title = "Table DATA 2.1\nSummary Statistics of Internal Characteristics ",
          header = FALSE,
          single.row = TRUE)


## 2.2 Feature: Amenities/Public Services
amenity.feature.list <- all.feature.list %>%
  dplyr:: select(starts_with('landmark_'),starts_with('mall_'),starts_with('hospital_'),
                 starts_with('college_'),starts_with('school_'),starts_with('hotel_'),
                 pctTotalVacant, pctRenterOccupied, pctWhite,pctPoverty, MedRent ,MedHHInc)

stargazer(amenity.feature.list, type = "html", 
          title = "Table DATA 2.2\nSummary Statistics of Amenities/Public Services ",
          header = FALSE,
          single.row = TRUE)

## 2.3 Spatial Structure: Neighborhood
spatial.feature.list <- all.feature.list %>%
  dplyr:: select(LABEL) 

as.data.frame(table(spatial.feature.list)) %>%
  rename(`Neighborhood List` = spatial.feature.list,
         Number =Freq) %>%
  kable(caption = 'Table DATA 2.3\nSpatial Feature: Neighborhood') %>%
  kable_styling("striped", full_width = F)

```


### 2.3 Correlation Matrix
```{r Correlation Matrix, echo=TRUE, message=FALSE, warning=FALSE}
# DATA - 3 Correlation Matrix
miami.numericVars.internal <- 
  select_if(internal.feature.list, is.numeric) %>% 
  na.omit()

miami.numericVars.amenity <- 
  select_if(amenity.feature.list, is.numeric) %>% 
  na.omit()

ggcorrplot(
  round(cor(miami.numericVars.internal), 1), 
  p.mat = cor_pmat(miami.numericVars.internal),
  colors = c('#05A167', "white", '#6897BB'),
  type="lower",
  insig = "blank") +  
  labs(title = "Correlation across numeric variables (internal features)\n",
       caption = 'Figure DATA 3.1') +
  plotTheme()

ggcorrplot(
  round(cor(miami.numericVars.amenity), 1), 
  p.mat = cor_pmat(miami.numericVars.amenity),
  colors = c('#05A167', "white", '#6897BB'),
  type="lower",
  insig = "blank") +  
  labs(title = "Correlation across numeric variables (amenity/public service features)\n",
       caption = 'Figure DATA 3.2') +
  plotTheme()

```


### 2.4 Home Price Correlation Scatterplots
```{r Scatter plot, echo=TRUE, message=FALSE, warning=FALSE}
# DATA - 4 Scatter plot  
## 1. Landmark Correlation 
price.landmark.plot <-
  st_drop_geometry(miami.training) %>% 
  dplyr::select(SalePrice, starts_with("landmark_")) %>%
  filter(SalePrice <= 1000000) %>%
  gather(Variable, Value, -SalePrice) %>% 
  ggplot(aes(Value, SalePrice)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#6897BB") +
  facet_wrap(~Variable, nrow = 1, scales = "free") +
  labs(title = "Price as a function of continuous variables\n",
       caption = 'Figure DATA xxx') +
  plotTheme()

price.landmark.plot

## 2. Shopping Mall Correlation 
price.mall.plot <-
  st_drop_geometry(miami.training) %>% 
  dplyr::select(SalePrice, starts_with("mall_")) %>%
  filter(SalePrice <= 1000000) %>%
  gather(Variable, Value, -SalePrice) %>% 
  ggplot(aes(Value, SalePrice)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#6897BB") +
  facet_wrap(~Variable, nrow = 1, scales = "free") +
  labs(title = "Price as a function of continuous variables\n",
       caption = 'Figure DATA xxx') +
  plotTheme()

price.mall.plot

## 3. Sexual Assaults Correlation 
price.sexual.plot <-
  st_drop_geometry(miami.training) %>% 
  dplyr::select(SalePrice, starts_with("sexual_")) %>%
  filter(SalePrice <= 1000000) %>%
  gather(Variable, Value, -SalePrice) %>% 
  ggplot(aes(Value, SalePrice)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#6897BB") +
  facet_wrap(~Variable, nrow = 1, scales = "free") +
  labs(title = "Price as a function of continuous variables\n",
       caption = 'Figure DATA xxx') +
  plotTheme()

price.sexual.plot


## 4. House Internal characteristics
price.house.plot <- 
  st_drop_geometry(miami.training) %>% 
  dplyr::select(SalePrice, pctTotalVacant, AdjustedSqFt, LotSize) %>%
  filter(SalePrice <= 1000000) %>%
  gather(Variable, Value, -SalePrice) %>% 
  ggplot(aes(Value, SalePrice)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#6897BB") +
  facet_wrap(~Variable, ncol = 3, scales = "free") +
  plotTheme() + 
  labs(title = "Price as a function of house internal characteristics variables\n",
       caption = 'Figure DATA xxx') 

price.house.plot

```

### 2.5 Map of Sale Price per Square Feet
```{r Sale Price Map, echo=TRUE, message=FALSE, warning=FALSE}
# DATA - 5 Map of Sale Price per Square Feet
# 
miami.training <- miami.training %>%
  mutate(priceFt = SalePrice/LivingSqFt)

map.priceFt <- ggplot() + 
  geom_sf(data = st_union(miami.base),fill = 'grey') +
  geom_sf(data = st_centroid(miami.training),aes(color = q5(priceFt)),size = .5) +
  scale_color_manual(values = palette,
                     labels = qBr(miami.training,'priceFt'),
                     name = "Price/ft^2") +
  labs(title = 'Sale Price Per Square Foot\n',
       subtitle = '',
       caption = 'Figure DATA 5') +
  mapTheme() + 
  plotTheme()

map.priceFt

```

### 2.6 Maps of Independent Variables
#### 2.6.1 Map of Landmark
```{r landmarkmap, echo=TRUE, message=FALSE, warning=FALSE}
# DATA - 6 

## 1. Landmark
plot.landmark <- ggplot() + 
  geom_sf(data = miami.base, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(miami.landmark.sf)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#05A167", high = "#6897BB", name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of landmark Assaults, Miami\n",
       caption = 'Figure DATA xxx') +
  mapTheme()

plot.landmark
```

#### 2.6.2 Map of Shopping Mall
```{r shoppingmap, echo=TRUE, message=FALSE, warning=FALSE}
## 2. Shopping Mall
plot.mall <- ggplot() + 
  geom_sf(data = miami.base, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(miami.mall.sf)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#05A167", high = "#6897BB", name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Shopping Malls, Miami\n",
       caption = 'Figure DATA xxx') +
  mapTheme()

plot.mall
```

#### 2.6.3 Map of Crime - Sexual Assaults
```{r crimemap, echo=TRUE, message=FALSE, warning=FALSE}
## 3. Sexual Assaults 

plot.sexual <- ggplot() + 
  geom_sf(data = miami.base, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(miami.sexual.sf)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#05A167", high = "#6897BB", name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Sexual Assaults, Miami\n",
       caption = 'Figure DATA xxx') +
  mapTheme()

plot.sexual

```

## III. Methods - Model Building

```{r Model, echo=TRUE, message=FALSE, warning=FALSE, results ='asis'}
## First model: All internal features, amenity features
M1 <- lm(SalePrice ~ ., data = st_drop_geometry(miami.training) %>% 
           dplyr::select(colnames(internal.feature.list), 
                         colnames(amenity.feature.list)
))

stargazer(M1, type = "html", 
          title = "Summary Statistics of Model 1 ",
          header = FALSE,
          single.row = TRUE)

## Second model: 
M2_1 <- lm(SalePrice ~ ., data = st_drop_geometry(miami.training) %>% 
           dplyr::select(# internal features
                         SalePrice, Land, Bldg, Assessed, County.Taxable,
                         AdjustedSqFt, LotSize, Bed,Bath, ActualSqFt, 
                         # amenity features
                         landmark_nn2, landmark_nn3, college_nn1,
                         school_nn3, school_nn4, school_nn5,
                         pctTotalVacant, pctRenterOccupied, MedRent, MedHHInc) 
)

stargazer(M2_1, type = "html", 
          title = "Summary Statistics of Model 2.1 ",
          header = FALSE,
          single.row = TRUE)

M2_2 <- lm(SalePrice ~ ., data = st_drop_geometry(miami.training) %>% 
             dplyr::select(# internal features
               SalePrice, Land, Bldg, Assessed, County.Taxable,
               AdjustedSqFt, LotSize, Bed,Bath, ActualSqFt, 
               # amenity features
               college_nn1,
               school_nn3, school_nn4, school_nn5,
               pctTotalVacant, MedRent, MedHHInc) 
)

stargazer(M2_2, type = "html", 
          title = "Summary Statistics of Model 2.2 ",
          header = FALSE,
          single.row = TRUE)

## Below is our baseline regression model
m2 <- M2_2

```

## IV. Result

### 4.1 Split dataset

```{r Split, echo=TRUE, message=FALSE, warning=FALSE}
# 1. Split dataset
set.seed(31357)

# get index for training sample
inTrain <- caret::createDataPartition(
  y = paste(miami.training$LABEL,miami.training$SalePrice ), 
  p = .60, list = FALSE)
# split data into training and test
model.miami.training <- miami.training[inTrain,] 
model.miami.test <- miami.training[-inTrain,] 

```

### 4.2 Training Set Summary Results

```{r Training, echo=TRUE, message=FALSE, warning=FALSE, results ='asis'}
# 2. Table of model summary on my training set
m2.training <- lm(SalePrice ~ ., data = st_drop_geometry(model.miami.training) %>% 
           dplyr::select(# internal features
             SalePrice, Land, Bldg, Assessed, County.Taxable,
             AdjustedSqFt, LotSize, Bed,Bath, ActualSqFt, 
             # amenity features
             college_nn1,
             school_nn3, school_nn4, school_nn5,
             pctTotalVacant, MedRent, MedHHInc
           ) 
)


stargazer(m2.training, type = "html", 
          title = "Summary Statistics of Baseline Model on Training Set",
          header = FALSE,
          single.row = TRUE)

broom::glance(m2.training) %>%
  kable(caption = 'Summary of Model Evaluation Parameters') %>%
  kable_styling("striped", full_width = F)

```
### 4.3 MAE, MAPE on Test Set

```{r MAE, echo=TRUE, message=FALSE, warning=FALSE}
# 3. MAE, MAPE on my test set

m2.miami.test <-
  model.miami.test %>% 
  mutate(Regression = "Baseline Regression",
         SalePrice.Predict = predict(m2.training, newdata = model.miami.test),
         SalePrice.Error = SalePrice - SalePrice.Predict,
         SalePrice.AbsError = abs(SalePrice - SalePrice.Predict),
         SalePrice.APE = (abs(SalePrice - SalePrice.Predict)) / SalePrice) %>%
  filter(SalePrice < 5000000) 
  
test_GoodnessFit<-as.data.frame(cbind(mean(m2.miami.test$SalePrice.AbsError,na.rm = T),
                                      mean(m2.miami.test$SalePrice.APE,na.rm = T)))
colnames(test_GoodnessFit)<-c("Mean Absolute Errors (MAE)","MAPE")
kable(test_GoodnessFit,
      caption = 'Table RESULT 3. MAE and MAPE for Test Set') %>%
  kable_styling("striped", full_width = F)

```
### 4.4 Cross-Validation
```{r Cross-Validation, echo=TRUE, message=FALSE, warning=FALSE}
# 4. Cross-Validation Test on Model 2
## 4.1 use caret package cross-validation method
fitControl <- caret:: trainControl(method = "cv", 
                                   number = 100,
                                   savePredictions = TRUE)

set.seed(717)

m2.cv <- 
  caret::train(SalePrice ~ ., data = st_drop_geometry(miami.training) %>% 
                 dplyr::select(# internal features
                   SalePrice, Land, Bldg, Assessed, County.Taxable,
                   AdjustedSqFt, LotSize, Bed,Bath, ActualSqFt, 
                   # amenity features
                   college_nn1,
                   school_nn3, school_nn4, school_nn5,
                   pctTotalVacant, MedRent, MedHHInc) ,
               method = "lm", 
               trControl = fitControl, 
               na.action = na.pass)

## 4.2 show RMSE, MAE, R^2
kable(m2.cv$resample,
          caption = 'Table RESULT 4. Cross-validation Test: Summary of RMSE, R Squared and MAE') %>%
  kable_styling("striped", full_width = F)

m2.cv$resample %>% 
  pivot_longer(-Resample) %>% 
  mutate(name = as.factor(name)) %>% 
  ggplot(., aes(x = name, y = value, color = name)) +
  geom_jitter(width = 0.1) +
  facet_wrap(~name, ncol = 3, scales = "free") +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = 'Cross_validation Test: Distribution of MAE, RMSE, R Squared\n',
       caption = "Figure RESULT 4.1") +
  plotTheme()


ggplot(data = m2.cv$resample) +
  geom_histogram(aes(x = m2.cv$resample$MAE), fill = 'orange') +
  labs(title="Distribution of Cross-validation MAE\n",
       subtitle = "K = 100",
       caption = "Figure RESULT 4.2") +
  xlab('MAE of Model 2') +
  ylab('Count') +
  plotTheme()
  
# If the model generalized well, the distribution of errors would cluster tightly together. 
# Instead, this range of errors suggests the model predicts inconsistently, and would likely be unreliable for predicting houses that have not recently sold.

```

### 4.5 Predicted Prices Plot

```{r Predicted Price, echo=TRUE, message=FALSE, warning=FALSE}
# 5. Plot predicted prices as a function of observed prices
preds.train <- data.frame(pred   = predict(m2.training, model.miami.training),
                          actual = model.miami.training$SalePrice,
                          source = "training data")
preds.test  <- data.frame(pred   = predict(m2.training,newdata = model.miami.test),
                          actual = model.miami.test$SalePrice,
                          source = "test data")
preds <- rbind(preds.train, preds.test)

## Overall Plotting
ggplot(preds, aes(x = actual, y = pred, color = source)) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", ) +
  geom_abline(color = "orange") +
  coord_equal() +
  theme_bw() +
  labs(title = "Predicted Prices as a Function of Observed Prices\n",
       caption = 'Figure RESULT 5.1',
       x = "Observed Prices",
       y = "Predicted Prices") +
  theme(
    legend.position = "none"
  ) +
  plotTheme()

ggplot(preds, aes(x = actual, y = pred, color = source)) +
  geom_point() +
  geom_smooth(method = "lm", color = "green") +
  geom_abline(color = "orange") +
  coord_equal() +
  theme_bw() +
  facet_wrap(~source, ncol = 2) +
  labs(title = "Predicted Prices as a Function of Observed Prices, by Sets\n",
       caption = 'Figure RESULT 5.2',
       x = "Observed Prices",
       y = "Predicted Prices") +
  theme(
    legend.position = "none"
  ) +
  plotTheme()
```

### 4.6 Map of Test Set Residuals
```{r Residual, echo=TRUE, message=FALSE, warning=FALSE}
# 6. Residual Exploration
## 6.1 A map of residuals for test set
map.res.test <- ggplot() + 
  geom_sf(data = st_union(miami.base),fill = 'grey') +
  geom_sf(data = m2.miami.test,aes(color = q5(SalePrice.AbsError)), show.legend = "point",size = 1) +
  scale_color_manual(values = palette,
                     labels = qBr(m2.miami.test,'SalePrice.AbsError'),
                     name = "Residuals") +
  labs(title = 'Map of residuals for test set\n',
       caption = 'Figure 6.1') +
  mapTheme() + 
  plotTheme()
map.res.test

## 6.2 Spatial lag in errors: average errors in five nearest neighbors
k_nearest_neighbors = 5

# average errors in five nearest neighbors
coords.test <- st_centroid(st_geometry(m2.miami.test), of_largest_polygon=TRUE)
neighborList.test <- knn2nb(knearneigh(coords.test, k_nearest_neighbors))
spatialWeights.test <- nb2listw(neighborList.test, style="W") 
m2.miami.test$lagPriceError <- lag.listw(spatialWeights.test,m2.miami.test$SalePrice.Error)

error.spatial.lag <- ggplot(m2.miami.test, aes(x = lagPriceError, y = SalePrice)) +
  geom_point(colour = "#6897BB") +
  geom_smooth(method = "lm", se = FALSE, colour = "#05A167") +
  labs(title = "Error on Test Set as a function of the Spatial Lag of Price",
       caption = "Figure RESULT 6.2",
       x = "Spatial lag of errors (Mean error of 5 nearest neighbors)",
       y = "Sale Price") +
  plotTheme()

error.spatial.lag

## 6.3 Moran's I Test
moranTest <- moran.mc(m2.miami.test$SalePrice.AbsError, 
                      spatialWeights.test, nsim = 999)

moran.plot <- ggplot(as.data.frame(moranTest$res[c(1:999)]), aes(moranTest$res[c(1:999)])) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(aes(xintercept = moranTest$statistic), colour = "#6897BB",size=1) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title="Observed and permuted Moran's I",
       subtitle= "Observed Moran's I in orange",
       x="Moran's I",
       y="Count",
       caption="Figure Result 6.3") +
  plotTheme()

moran.plot



```

### 4.7 Map of Predicted Prices on the Whole Dataset

```{r Whole map, echo=TRUE, message=FALSE, warning=FALSE}
# 7. Map of predicted values for the whole dataset
miami.sf$Predicted.Price<- predict(m2.training, newdata = miami.sf)

map.predicted.price <- ggplot() + 
  geom_sf(data = st_union(miami.base),fill = 'grey') +
  geom_sf(data = st_centroid(miami.sf),aes(color = q5(Predicted.Price)),size = .5) +
  scale_color_manual(values = palette,
                     labels = qBr(miami.sf,'Predicted.Price'),
                     name = "Price, $") +
  labs(title = 'Map of Predicted Sale Price\n',
       subtitle = '',
       caption = 'Figure RESULT 7') +
  mapTheme() + 
  plotTheme()

map.predicted.price
```

### 4.8 Map of Mean Absolute Percentage Error (MAPE) by Neighborhood

```{r MAPE, echo=TRUE, message=FALSE, warning=FALSE}
# 8. Map of MAPE by neighborhood
## 8.1 Statistic summary by neighborhood group
nhood_sum <- m2.miami.test %>% 
  group_by(LABEL) %>%
  summarize(meanPrice = mean(SalePrice, na.rm = T),
            meanPrediction = mean(SalePrice.Predict, na.rm = T),
            meanMAE = mean(SalePrice.AbsError, na.rm = T),
            meanMAPE = mean(SalePrice.APE, na.rm = T)) 

nhood_sum %>%
  st_drop_geometry %>%
  arrange(desc(meanMAPE)) %>% 
  kable(caption = 'Table RESULT 8. Map of MAPE by Neighborhood') %>% 
  kable_styling("striped", full_width = F)

## 8.2 Map of MAPE by neighborhood

miami.neigh.sum <- miami.neigh %>%
  left_join(st_drop_geometry(nhood_sum), by = "LABEL") %>%
  mutate(meanMAPE.expanded = meanMAPE * 100)
  
            
map.MAPE.nhood <- ggplot() + 
  geom_sf(data = miami.neigh.sum,
          aes(fill = q5(meanMAPE.expanded))) + 
  scale_fill_manual(values = palette,
                     labels = qBr(miami.neigh.sum,'meanMAPE.expanded'),
                     name = "MAPE\n(timed by 100 for better display)") +
  labs(title = 'Map of MAPE on Test Set, by Neighborhood\n',
       subtitle = '',
       caption = 'Figure RESULT 8') +
  mapTheme() + 
  plotTheme()

map.MAPE.nhood

```

### 4.9 Scatter plot of MAPE by neighborhood as a function of mean price by neighborhood

```{r MAPE Scatter, echo=TRUE, message=FALSE, warning=FALSE}
# 9. Scatter plot of MAPE by neighborhood as a function of mean price by neighborhood

MAPE.nhood.plot <-ggplot()+
  geom_point(data = nhood_sum, aes(x = meanPrice, y = meanMAPE)) +
  stat_smooth(data = nhood_sum, aes(x = meanPrice, y = meanMAPE), method = "loess", se = FALSE, size = 1, colour="red") + 
  labs(title = "MAPE by Neighborhood as a Function of Mean Price by Neighborhood\n",
       caption = 'Figure RESULT 9') +
  xlab('Mean Price by Neighborhood') +
  ylab('Mean MAPE by Neighborhood') +
  plotTheme()

MAPE.nhood.plot
```

### 4.10 Generalizability

```{r general, echo=TRUE, message=FALSE, warning=FALSE}

# 10. Model generalizability on census data
## 10.1 Group Plotting
census <-   miami.training %>% # miami.training contains all features we used including census data and geometry information
   mutate(raceContext = ifelse(pctWhite > .5, "Majority White", "Majority Non-White"),
          incomeContext = ifelse(MedHHInc > mean(MedHHInc,na.rm = T), "High Income", "Low Income"),
          vacantContext = ifelse(pctTotalVacant > .5, "Majority Vacant", "Majority Non-Vacant"),
          pctRenterContext = ifelse(pctPoverty > .5, "Majority Renter Occupied", "Majority Non-Renter Occupied")) %>%
  select(raceContext,incomeContext,vacantContext,pctRenterContext)


census <- census %>%
  st_transform(st_crs(tracts17))


grid.arrange(ncol = 2,
             ggplot() + 
               geom_sf(data = st_union(miami.base), fill = 'grey') + 
               geom_sf(data = na.omit(census), aes(color = raceContext),
                       show.legend = "point", size= 1) +
               scale_color_manual(values = c("#05A167", "#6897BB"), name="Race Context") +
               labs(title = "Race Context of House\n") +
               mapTheme() + 
               theme(legend.position="bottom") +
               plotTheme(),
             
             ggplot() + 
               geom_sf(data = st_union(miami.base), fill = 'grey') + 
               geom_sf(data = na.omit(census), aes(color = incomeContext),
                       show.legend = "point", size= 1) +
               scale_fill_manual(values = c("#05A167", "#6897BB"), name="Income Context") +
               labs(title = "Income Context of House\n") +
               mapTheme() + 
               theme(legend.position="bottom") +
                plotTheme())

grid.arrange(ncol = 2,
             ggplot() + 
               geom_sf(data = st_union(miami.base), fill = 'grey') + 
               geom_sf(data = na.omit(census), aes(color = vacantContext),
                       show.legend = "point", size= 1) +
               scale_color_manual(values = c("#05A167", "#6897bb"), name="Vacant Context") +
               labs(title = "Vacant Context of House\n") +
               mapTheme() + 
               theme(legend.position="bottom") +
               plotTheme(),
             ggplot() + 
               geom_sf(data = st_union(miami.base), fill = 'grey') + 
               geom_sf(data = na.omit(census), aes(color = pctRenterContext),
                       show.legend = "point", size= 1) +
               scale_fill_manual(values = c("#05A167", "#6897BB"), name="Renter Proportion Context") +
               labs(title = "Renter Proportion Context of House\n") +
               mapTheme() + 
               theme(legend.position="bottom") +
                plotTheme())




```

```{r Fixed Effect, echo=TRUE, message=FALSE, warning=FALSE}
### Based on the group distribution, we decided to select race group and income group

## 10.2 Neighborhood Fixed Effect
### 10.2.1 Model building
m2.nhood <- lm(SalePrice ~ ., data = as.data.frame(model.miami.training) %>% 
     dplyr::select(# spatial features
                   LABEL,
                   # internal features
                   SalePrice, Land, Bldg, Assessed, County.Taxable,
                   AdjustedSqFt, LotSize, Bed,Bath, ActualSqFt, 
                   # amenity features
                   college_nn1,
                   school_nn3, school_nn4, school_nn5,
                   pctTotalVacant, MedRent, MedHHInc) 
)

m2.miami.test.nhood <-
  model.miami.test %>% 
  mutate(Regression = "Neighborhood Effects",
         SalePrice.Predict = predict(m2.nhood, model.miami.test),
         SalePrice.Error = SalePrice - SalePrice.Predict,
         SalePrice.AbsError = abs(SalePrice - SalePrice.Predict),
         SalePrice.APE = (abs(SalePrice - SalePrice.Predict)) / SalePrice) %>%
  filter(SalePrice < 5000000)

### 10.2.2. Model combination and comparison
bothRegressions <- 
  rbind(
    dplyr::select(m2.miami.test, starts_with("SalePrice"), Regression, LABEL) ,
    dplyr::select(m2.miami.test.nhood, starts_with("SalePrice"), Regression, LABEL) )

st_drop_geometry(bothRegressions) %>%
  gather(Variable, Value, -Regression, -LABEL) %>%
  filter(Variable == "SalePrice.AbsError" | Variable == "SalePrice.APE") %>%
  group_by(Regression, Variable) %>%
  summarize(meanValue = mean(Value, na.rm = T)) %>%
  spread(Variable, meanValue) %>%
  kable(caption = 'Table RESULT 10.1.\nSummary of MAE, and MAPE by Regression Models') %>%
  kable_styling("striped", full_width = F) %>%
  row_spec(1, color = "black", background = "#b2eed7") %>%
  row_spec(2, color = "black", background = '#6897BB')

### 10.2.3. Model coefficents for each Neighborhood
tidy(m2.nhood) %>% 
  filter(str_detect(term, "LABEL")) %>% 
  kable(caption = " RESULT 10.2.\nModel Coefficents for Each Neighborhood") %>% 
  kable_styling("striped", full_width = F)

### 10.2.4 Generalizability in Different Groups
census <-   miami.training %>% # miami.training contains all features we used including census data and geometry information
   mutate(raceContext = ifelse(pctWhite > .5, "Majority White", "Majority Non-White"),
          incomeContext = ifelse(MedHHInc > mean(MedHHInc,na.rm = T), "High Income", "Low Income"),
          vacantContext = ifelse(pctTotalVacant > .5, "Majority Vacant", "Majority Non-Vacant"),
          pctRenterContext = ifelse(pctPoverty > .5, "Majority Renter Occupied", "Majority Non-Renter Occupied")) %>%
  select(raceContext,incomeContext,vacantContext,pctRenterContext)

race.group <- st_join(bothRegressions, census) %>% 
  group_by(Regression, raceContext) %>%
  summarize(mean.MAPE = scales::percent(mean(SalePrice.APE, na.rm = T))) %>%
  st_drop_geometry() %>%
  spread(raceContext, mean.MAPE) %>%
  kable(caption = "Table  RESULT 10.3.\nTest set MAPE by Neighborhood Racial Context") %>% 
  kable_styling("striped", full_width = F)

race.group

income.group <- st_join(bothRegressions, census) %>% 
  group_by(Regression, incomeContext) %>%
  summarize(mean.MAPE = scales::percent(mean(SalePrice.APE, na.rm = T))) %>%
  st_drop_geometry() %>%
  spread(incomeContext, mean.MAPE) %>%
  kable(caption  = " RESULT 10.4.\nTest set MAPE by Neighborhood Income Context") %>% 
  kable_styling("striped", full_width = F)

income.group

```
## V. Discussion
Is this an effective model? 
What were some of the more interesting variables?  
How much of the variation in prices could you predict? 
Describe the more important features? 
Describe the error in your predictions?  
According to your maps, could you account the spatial variation in prices?  
Where did the model predict particularly well? Poorly? Why do you think this might be?


## VI. Conclusion
Would you recommend your model to Zillow? Why or why not? 
How might you improve this model?  
	Incorporate restaurant and bar features. Since Miami is famous for night life and party, restaurant and bar location might be relevant to the house price. However, there is no open data available for Miami restaurants. 


